Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-03-25 ForegroundSimulations/ms.tex
Make updates suggested by Aaron Parsons
From bbb5b82 (fgsim):
====================

In eor.bib,

(*) Add WMAP9 and Planck references for cosmological parameters

In ms.tex,

(*) Make updates suggested by Aaron Parsons

Unaccounted systematics from foregrounds and instruments prevent current experiments from detecting redshifted 21~cm signal from the Epoch of Reionization (EoR). We identify spectral structures (chromaticity) in the antenna power pattern as one such systematic that significantly affects all EoR experiments. Increase in chromaticity of the antenna beam causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated design specifications on these reflections to prevent further EoR signal degradation. We use the Hydrogen Epoch of Reionization Array (HERA) as an example. Using a simple and yet high dynamic range spectral weighting scheme, we demonstrate that EoR detection using HERA will not be impeded by such spectral systematics even in a conservative scenario that does not perform removal of foregrounds. All baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

Numerous experiments using low frequency radio telescopes targeting the redshifted 21~cm line from the spin-flip transition of H{\sc i} in this epoch have become operational such as the Murchison Widefield Array \citep[MWA;][]{lon09,bow13,tin13}, the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, the Low Frequency Array \citep[LOFAR;][]{van13} and the Giant Metrewave Radio Telescope EoR experiment \citep[GMRT;][]{pac13}. Many of these instruments have sufficient sensitivity for a statistical detection of the EoR signal by estimating the spatial power spectrum of the redshifted H{\sc i} spin temperature fluctuations \citep{bea13,thy13}. These instruments are intended to be precursors and pathfinders to the next generation of low frequency radio observatories such as the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} \citep[HERA;][]{deb16} and the Square Kilometre Array\footnote{\url{https://www.skatelescope.org/}} (SKA). These next-generation instruments will advance the capability from a mere statistical detection of the signal to a direct three-dimensional tomographic imaging of H{\sc i} from the EoR.

@@

Despite attempts to exploit this knowledge about differences in spectral and spatial symmetry, first-generation experiments have at best only led to weak constraints on reionization models \citep{ali15,dil15,pob15}. This is because their sensitivity is primarily limited by instrumental and foreground systematics. The extreme dynamic range required to suppress foreground and instrument systematics demands high precision modeling of foregrounds as observed by modern wide-field instruments.

@@

In this paper, we explore a phenomenon that inherently extends foreground power beyond the {\it wedge}, namely, that arising from the chromaticity of the antenna power pattern. While highlighting the importance of such instrumental systematics on future EoR experiments, we demonstrate that the current HERA instrument design will exceed these specifications to detect the EoR signal with high significance. This will hold true when data is limited by foregrounds and systematics even in the most conservative scenario which involves neither foreground subtraction nor optimal analysis techniques but only relies on a simple spectral weighting scheme.

@@

\citet{deb16} describe the HERA instrument in detail. We provide a summary of the instrument currently under construction on the South African SKA site.

@@

HERA will deploy 14~m fixed zenith-pointing parabolic dishes that strike an optimal balance between sensitivity and systematics. The large collecting area of a HERA element yields $\approx 5$ times the sensitivity of an MWA tile and more than 20 times that of a PAPER element.

@@

The delay spectrum has a close resemblance to cosmological H{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P(\boldsymbol{b},k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\,f_{21}H_0\,E(z)}{c(1+z)^2}\,\tau, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $\boldsymbol{k}_\perp$ and $k_\parallel$ are the transverse (on the sky) and line-of-sight (into the sky) wavenumbers respectively, $f_{21}$ is the rest-frame frequency of the 21~cm spin-flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. $P(\boldsymbol{b},k_\parallel)$ is in units of K$^2 (h^{-1}$~Mpc$)^3$. In this paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$. 
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f,
\end{align}
is related to the cosmic volume probed by the instrument \citet{par14}.

@@

The equation for delay spectrum describes the mapping between sky location of a foreground object and delay. The chromatic nature (variation with frequency) of the antenna power pattern appears multiplicatively in the frequency axis and thus results in convolving its delay response with that of the foregrounds and the geometric phases measured by the baseline vector. We discuss the resulting spillover of foreground power beyond the horizon delay limits.

@@

Since our aim in this paper is to quantify EoR-foreground power ratio without employing any sophisticated {\it foreground removal} techniques, we turn our attention to {\it foreground avoidance} strategy that employs spectral weighting technique. Specific choices for spectral weighting function, $W(f)$, have been found to be effective in reducing foreground contamination by many orders of magnitude \citep{thy13} and is regularly used in EoR data analysis \citep{par12a}. For instance, a {\it Blackman-Harris} function \citep{har78} has a dynamic range of $\sim$100 -- 120~dB in delay power spectrum and a reduced effective bandwidth, with:
\begin{align}\label{eqn:Beff}
  \epsilon &= \frac{1}{B} \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f, \\
  \textrm{and,}\quad B_\textrm{eff} &= \epsilon\,B,
  % B_\textrm{eff} &= \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f.
\end{align}
where, $B$ denotes the end-to-end range of the observing band, $1-\epsilon$ is the loss in overall spectral sensitivity and $B_\textrm{eff}$ is the effective bandwidth. For a {\it Blackman-Harris} window function denoted by $W_\textrm{BH}(f)$, $\epsilon \approx 50\%$. % Hence, the window spans a total of $2N$ frequency channels each of width $\Delta f$ to have $B_\textrm{eff} \approx N\,\Delta f$.

@@

Fig.~\ref{fig:window-functions} shows the {\it Blackman-Harris} window (gray) and  the window in equation~\ref{eqn:bhw2} (black) in frequency (top) and their power responses in delay (bottom). In its Fourier domain, this window has a response obtained by squaring the {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB with $\epsilon\approx 42\%$.

@@

This clearly demonstrates that higher the beam chromaticity, the farther the foreground contamination inherently extends along $k_\parallel$. Thus, the chromaticity of antenna beam needs to be controlled in EoR experiments to keep the foreground systematics sufficiently low.

@@

Reflections shift the measured foreground power to higher modes in $\tau$ (or in $k_\parallel$) and thus cause further contamination in these critical higher $k_\parallel$-modes. These delay shifts introduce ripples in the spectrum. The net chromaticity in the measurements is the product of spectral structure arising out of the intrinsic nature of foreground emission, the baseline- and position-dependent frequency structure of the geometric phases, the spectral features in the antenna power pattern besides any other spectral structures in the instrument unaccounted for. In the Fourier domain these factors have a convolving effect. In order to isolate the requirements on beam chromaticity from other sources of chromaticity, we start with an antenna beam in which those spectral structures are absent but the intrinsic chromaticity of foregrounds and the geometric phases of baselines are included.

We define the required attenuation on the reflected foreground power as the ratio: 
\begin{align}\label{eqn:attenuation}
  \Gamma_{k_\textrm{p}}(\boldsymbol{b},\tau) \geq \max_{|k_\parallel|>k_\textrm{p}}\left\{\frac{\left\langle P_\textrm{FG}(\boldsymbol{b},k_\parallel - \frac{\dif k_\parallel}{\dif \tau}\,\tau)\right\rangle}{P_\textrm{H{\sc i}}(|\boldsymbol{b}|,k_\parallel)}\right\}^{1/2},
\end{align}
where, $\tau$ is the delay caused by reflections, and $\dif k_\parallel/\dif \tau$ is the {\it jacobian} in the transformation of $\tau$ to $k_\parallel$ (see equation~\ref{eqn:kprll-delay}). $P_\textrm{H{\sc i}}(|\boldsymbol{b}|,k_\parallel)$ is the EoR H{\sc i} delay power spectrum. $P_\textrm{FG}(\boldsymbol{b},k_\parallel)$ is the full-band foreground delay power spectrum. The angular brackets denote averaging over baseline vectors of the specified length, and a local sidereal time (LST) range. Thus, $\Gamma_{k_\textrm{p}}(\tau)$ is determined by the requirement that the reflected foreground power obtained by shifting in delay lies below the EoR H{\sc i} signal power in line-of-sight spatial scales of interest, $|k_\parallel|>k_\textrm{p}$.

@@

For reflections caused by dish-feed assembly, we use $P_\textrm{FG}(k_\parallel)$ obtained with the achromatic beam model on 14.6~m baselines in a 0--12 hr LST range. Fig.~\ref{fig:fg-reflections-achrmbeam} shows $\Gamma_{k_\textrm{p}}(\tau)$ (in dB) for $k_\textrm{p}$ chosen to be 0.1~$h$~Mpc$^{-1}$ (solid), 0.15~$h$~Mpc$^{-1}$ (dashed), and 0.2~$h$~Mpc$^{-1}$ (dotted). These curves set an upper limit for the reflected foreground power to lie below the EoR H{\sc i} signal power as a function of delays. These reflections are primarily caused in the dish-feed assembly and to some extent in other subsystems not accounted for. It implies that if attenuated to levels that lie in the regions below the different shaded regions, such spectral systematics in the instrument will not hinder detection of EoR in those corresponding $k_\parallel$-modes of interest.

@@

The elbow-shaped turnover is a measure of the most severe requirement on attenuation of reflections. This depends sensitively on the choice of $k_p$. For instance, the attenuation required is $\gtrsim 54$~dB at $\sim 200$~ns for $k_\textrm{p}=0.1\,h$~Mpc$^{-1}$, $\gtrsim 56$~dB at $\sim 300$~ns for $k_\textrm{p}=0.15\,h$~Mpc$^{-1}$ and $\gtrsim 58$~dB at $\sim 400$~ns for $k_\textrm{p}=0.2\,h$~Mpc$^{-1}$. Measurements are underway \citep{pat16} to confirm that the return losses in the HERA design lie well within the regions excluded by the shaded regions.

@@

Similarly, for reflections arising out of structures and interfaces across different antennas, we use $P_\textrm{FG}(k_\parallel)$ obtained with the simulated chromatic and {\it Airy} antenna beams on 14.6~m baselines. These beams do not include spectral features from antenna-to-antenna reflections. Thus $\Gamma_{k_\textrm{p}}(\tau)$ provides lower limit for attenuation of antenna-to-antenna reflections. Fig.~\ref{fig:fg-reflections} is similar to Fig.~\ref{fig:fg-reflections-achrmbeam} with constraints for values of $k_\textrm{p}$ chosen to be 0.1~$h$~Mpc$^{-1}$ (left), 0.15~$h$~Mpc$^{-1}$ (middle), and 0.2~$h$~Mpc$^{-1}$ (right). Lower limits are estimated for {\it Airy} (dashed lines) and simulated chromatic beams (solid lines).

@@

Similar design limits on reflections on various antenna spacings were also studied. With increase in baseline length, the amplitude of foreground power decreases but the {\it foreground window} also widens. This means the amplitude of reflections required to be suppressed is lower but there is lesser room along delay axis before the reflected foreground modes shift to and contaminate the $k_\parallel$-modes of interest. As a result, the elbow-shaped design constraints shift upward and leftward. For instance, in the presence of dish-feed reflections, it will still be feasible to probe $k_\parallel > 0.15\,h$~Mpc$^{-1}$ with the 58.4~m baselines but not other larger scales.

@@

We note that the inevitable reflections in the antenna-feed assembly and between reflecting structures across multiple antennas are also significant factors that contaminate the cosmological signal in $k_\parallel$-modes in the {\it EoR window}. Using a novel approach, we provide a formalism to set cosmologically-motivated design requirements directly on the level of suppression required on these reflections in the instrument. % Ongoing reflectometry studies clearly indicate HERA will comfortably achieve the design limits prescribed here.

By accounting for these instrumental systematics, if data is not limited by thermal noise, we demonstrate that HERA should detect EoR with high significance with a simple spectral weighting even under a conservative analysis criterion that does not involve {\it foreground removal} strategies. All baselines of HERA-19 will not only detect but also help in filtering highly likely EoR models.

We are also investigating advanced strategies for foreground removal that will robustly eliminate these systematics further. More optimal analysis techniques and further design improvements will make HERA not only clear the obstacles limiting first-generation experiments but also become transformational for next generation of low-frequency cosmology experiments.

@@

We thank Zaki Ali, Adam Beardsley, and Daniel Jacobs for their valuable inputs during the preparation of this manuscript.

>>>>

Unaccounted systematics from foregrounds and instruments limit sensitivity of current experiments from detecting redshifted 21~cm signal from the Epoch of Reionization (EoR). Upcoming experiments are faced with a challenge to deliver more collecting area per antenna element without degrading the data with systematics. This paper and its companions show that dishes are viable for achieving this balance using the Hydrogen Epoch of Reionization Array (HERA) as an example. In this paper, we specifically identify spectral systematics associated with the antenna power pattern as a significant detriment affecting all EoR experiments. Chromaticity of the antenna beam causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated design specifications on these reflections to prevent further EoR signal degradation and show HERA will not be impeded by such spectral systematics. Using a simple and yet high dynamic range spectral weighting scheme, we demonstrate that even in a conservative scenario that does not perform removal of foregrounds, HERA will detect EoR signal in line-of-sight $k$-modes, $k_\parallel \gtrsim 0.2\,h$~Mpc$^{-1}$, with high significance. All baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

Numerous experiments using low frequency radio telescopes targeting the redshifted 21~cm line from the spin-flip transition of H{\sc i} in this epoch have become operational such as the Murchison Widefield Array \citep[MWA;][]{lon09,bow13,tin13}, the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, the Low Frequency Array \citep[LOFAR;][]{van13} and the Giant Metrewave Radio Telescope EoR experiment \citep[GMRT;][]{pac13}. Many of these instruments have sufficient sensitivity for a statistical detection of the EoR signal by estimating the spatial power spectrum of the redshifted H{\sc i} spin temperature fluctuations \citep{bea13,thy13}. These instruments are precursors and pathfinders to the next generation of low frequency radio observatories such as the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} \citep[HERA;][]{deb16} and the Square Kilometre Array\footnote{\url{https://www.skatelescope.org/}} (SKA). These next-generation instruments will advance the capability from a statistical detection of the signal to a direct three-dimensional tomographic imaging of H{\sc i} from the EoR.

@@

% Despite attempts to exploit this knowledge about differences in spectral and spatial symmetry, first-generation experiments have at best only led to weak constraints on reionization models \citep{ali15,dil15,pob15}. This is because their sensitivity is primarily limited by instrumental and foreground systematics. The extreme dynamic range required to suppress foreground and instrument systematics demands high precision modeling of foregrounds as observed by modern wide-field instruments.

Using this knowledge, experiments have begun constraining reionization models \citep{par14,ali15,pob15}. While this progress is encouraging, efforts are slowed by struggles to increase collecting area while contending with foreground and instrumental systematics. The extreme dynamic range required to suppress foreground and instrument systematics demands high precision modeling of foregrounds as observed by modern wide-field instruments.

@@

Along with companion papers \citep{ewa16,neb16,pat16}, we focus on determining whether larger antenna elements can be viable for EoR experiments. In this paper, we explore the impact of chromaticity in the antenna power pattern on extending foreground power beyond the {\it wedge}. While highlighting the importance of such instrumental systematics on future EoR experiments, we demonstrate that the current HERA instrument design will exceed these specifications to detect the EoR signal with high significance. This will hold true when data is limited by foregrounds and systematics even in the most conservative scenario which involves neither foreground subtraction nor optimal analysis techniques but only relies on a simple spectral weighting scheme.

@@

\citet{deb16} describe the instrument in detail while \citet{pob14} explored the range of astrophysical parameters that can be probed with HERA. We provide a summary of the instrument currently under construction on the South African SKA site.

@@

HERA is deploying 14~m fixed zenith-pointing parabolic dishes that strike an optimal balance between sensitivity and systematics \citep[][and this paper]{ewa16,neb16,pat16}. The large collecting area of a HERA element yields $\approx 5$ times the sensitivity of an MWA tile and more than 20 times that of a PAPER element.

@@

The delay spectrum has a close resemblance to cosmological H{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P(\boldsymbol{b},k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\,f_{21}H_0\,E(z)}{c(1+z)^2}\,\tau, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $\boldsymbol{k}_\perp$ and $k_\parallel$ are the transverse (on the sky) and line-of-sight (into the sky) wavenumbers respectively, $f_{21}$ is the rest-frame frequency of the 21~cm spin-flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. $P(\boldsymbol{b},k_\parallel)$ is in units of K$^2 (h^{-1}$~Mpc$)^3$. In this paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,h\,$km$\,$s$^{-1}\,$Mpc$^{-1}$ \citep{wmap9cosmo}. 
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f,
\end{align}
is related to the cosmic volume probed by the instrument \citep[see appendix of][]{par14}.

@@

The equation for delay spectrum defines the interplay between foregrounds and antenna power pattern, and the mapping between geometric phases of baselines and delays. We discuss the spillover of foreground power beyond the horizon delay limits caused by the chromatic nature of the antenna power pattern.

% The equation for delay spectrum describes the mapping between sky location of a foreground object and delay. The chromatic nature (variation with frequency) of the antenna power pattern appears multiplicatively in the frequency axis and thus results in convolving its delay response with that of the foregrounds and the geometric phases measured by the baseline vector. We discuss the resulting spillover of foreground power beyond the horizon delay limits.

@@

Since our aim in this paper is to quantify EoR-foreground power ratio without employing any sophisticated {\it foreground removal} techniques, we turn our attention to {\it foreground avoidance} strategy that employs spectral weighting technique. Specific choices for spectral weighting function, $W(f)$, have been found to be effective in reducing foreground contamination by many orders of magnitude \citep{thy13} and is regularly used in redshifted 21~cm data analysis \citep{par12a,par12b}. For instance, a {\it Blackman-Harris} function \citep{har78} has a dynamic range of $\sim$100 -- 120~dB in delay power spectrum and a reduced effective bandwidth, with:
\begin{align}\label{eqn:Beff}
  \epsilon &= \frac{1}{B} \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f, \\
  \textrm{and,}\quad B_\textrm{eff} &= \epsilon\,B,
  % B_\textrm{eff} &= \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f.
\end{align}
where, $B$ denotes the end-to-end range of the observing band, $1-\epsilon$ is the loss in overall spectral sensitivity and $B_\textrm{eff}$ is the effective bandwidth. The {\it noise equivalent width} is defined as $w_\textrm{ne}\equiv \sum_i |\widetilde{W}(\tau_i)|^2 / |\widetilde{W}(0)|^2$. For a {\it Blackman-Harris} window function denoted by $W_\textrm{BH}(f)$, $\epsilon \approx 50\%$ and $w_\textrm{ne} \approx 2$. % Hence, the window spans a total of $2N$ frequency channels each of width $\Delta f$ to have $B_\textrm{eff} \approx N\,\Delta f$.

@@

Fig.~\ref{fig:window-functions} shows the {\it Blackman-Harris} window (gray) and  the window in equation~\ref{eqn:bhw2} (black) in frequency (top) and their power responses in delay (bottom). In its Fourier domain, this window has a response obtained by squaring the {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB with $\epsilon\approx 42\%$ and $w_\textrm{ne} \approx 2.88$.

@@

This clearly demonstrates that with increasing beam chromaticity, the foreground contamination inherently extends farther along $k_\parallel$. Thus, the chromaticity of antenna beam needs to be controlled in EoR experiments to keep the foreground systematics sufficiently low.

@@

Reflections shift the measured foreground power to higher modes in $\tau$ (or in $k_\parallel$) and thus cause further contamination in these critical higher $k_\parallel$-modes. These delay shifts introduce ripples in the spectrum. The net chromaticity in the measurements is the product of spectral structure arising out of the intrinsic nature of foreground emission, $I(\hat{\boldsymbol{s}},f)$, the baseline- and position-dependent frequency structure of the geometric phases, $e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}$, the spectral features in the antenna power pattern, $A(\hat{\boldsymbol{s}},f)$, besides any other spectral structures in the instrument unaccounted for. In $P(\boldsymbol{b},k_\parallel)$ defined in the Fourier domain, these factors have a convolving effect.

From these factors, we isolate here the effects of chromaticity in the antenna power pattern caused by reflections in antenna structures and signal paths. And, we devise a cosmologically-motivated method to set design requirements on suppressing reflections in the instrument.

We define the required attenuation on the reflected foreground power as the ratio: 
\begin{align}\label{eqn:attenuation}
  \Gamma_{k_\textrm{p}}(\boldsymbol{b},\tau) \geq \max_{|k_\parallel|>k_\textrm{p}}\left\{\frac{\left\langle P_\textrm{FG}(\boldsymbol{b},k_\parallel - \frac{\dif k_\parallel}{\dif \tau}\,\tau)\right\rangle}{P_\textrm{H{\sc i}}(|\boldsymbol{b}|,k_\parallel)}\right\}^{1/2},
\end{align}
where, $\tau$ is the delay caused by reflections, and $\dif k_\parallel/\dif \tau$ is the {\it jacobian} in the transformation of $\tau$ to $k_\parallel$ (see equation~\ref{eqn:kprll-delay}). $P_\textrm{H{\sc i}}(|\boldsymbol{b}|,k_\parallel)$ is the EoR H{\sc i} delay power spectrum. $P_\textrm{FG}(\boldsymbol{b},k_\parallel)$ is the full-band foreground delay power spectrum. The angular brackets denote averaging over baseline vectors of the specified length, and a local sidereal time (LST) range.

We assume the interpretation that reflections cause additional spectral structures in the antenna power pattern in which they would be absent otherwise. Thus, we use an antenna power pattern in which those spectral structures from reflections which we are interested in addressing are absent but the intrinsic chromaticity of foregrounds and the geometric phases of baselines are included. $\Gamma_{k_\textrm{p}}(\tau)$ is determined by the requirement that the reflected foreground power so obtained, after shifting in delay, lies below the EoR H{\sc i} signal power in line-of-sight spatial scales of interest, $|k_\parallel|>k_\textrm{p}$. Thus, by disentangling spectral systematics inherent in the antenna power pattern from the overall EoR-foreground dynamic range, a design requirement that sets a limit on that systematic is obtained.

@@

For reflections caused by dish-feed assembly, we use $P_\textrm{FG}(k_\parallel)$ obtained with the achromatic beam model on 14.6~m baselines in a 0--12 hr LST range. In this instance, we assume dish-feed reflections predominantly imprint chromaticity in an otherwise achromatic antenna power pattern and $\Gamma_{k_\textrm{p}}(\tau)$ is interpreted as the design goal for attenuating these reflections. These reflections are primarily caused in the dish-feed assembly and to some extent in other subsystems not accounted for. Fig.~\ref{fig:fg-reflections-achrmbeam} shows $\Gamma_{k_\textrm{p}}(\tau)$ (in dB) for $k_\textrm{p}$ chosen to be 0.1~$h$~Mpc$^{-1}$ (solid), 0.15~$h$~Mpc$^{-1}$ (dashed), and 0.2~$h$~Mpc$^{-1}$ (dotted). These curves set an upper limit for the reflected foreground power to lie below the EoR H{\sc i} signal power as a function of delays. It implies that if attenuated to levels that lie in the regions below the different shaded regions, such spectral systematics in the instrument will not hinder detection of EoR in those corresponding $k_\parallel$-modes of interest.

@@

The elbow-shaped turnover is a measure of the most severe requirement on attenuation of reflections. This depends sensitively on the choice of $k_p$. For instance, the attenuation required is $\gtrsim 54$~dB at $\sim 200$~ns for $k_\textrm{p}=0.1\,h$~Mpc$^{-1}$, $\gtrsim 56$~dB at $\sim 300$~ns for $k_\textrm{p}=0.15\,h$~Mpc$^{-1}$ and $\gtrsim 58$~dB at $\sim 400$~ns for $k_\textrm{p}=0.2\,h$~Mpc$^{-1}$. Measurements are underway \citep{pat16} to confirm that the return losses in the HERA design lie within the regions excluded by the shaded regions.

@@

Similarly, for reflections arising out of structures and interfaces across different antennas, we use $P_\textrm{FG}(k_\parallel)$ obtained with the simulated chromatic and {\it Airy} antenna beams on 14.6~m baselines. The simulated chromatic pattern in fact includes explicit simulations of reflections between structures within an antenna subsystem while the {\it Airy} pattern is assumed to include these effects. But both these beams do not include spectral features from antenna-to-antenna reflections. Thus $\Gamma_{k_\textrm{p}}(\tau)$ provides lower limit for attenuation of such antenna-to-antenna reflections. Fig.~\ref{fig:fg-reflections} is similar to Fig.~\ref{fig:fg-reflections-achrmbeam} with constraints for values of $k_\textrm{p}$ chosen to be 0.1~$h$~Mpc$^{-1}$ (left), 0.15~$h$~Mpc$^{-1}$ (middle), and 0.2~$h$~Mpc$^{-1}$ (right). Lower limits are estimated for {\it Airy} (dashed lines) and simulated chromatic beams (solid lines).

@@

Similar design limits on reflections on various antenna spacings were also studied. With increase in baseline length, the amplitude of foreground power decreases but the {\it foreground window} also widens. This means the amplitude of reflections required to be suppressed is lower but there is lesser room along delay axis before the reflected foreground modes shift to and contaminate the $k_\parallel$-modes of interest. As a result, the elbow-shaped design constraints shift upward and leftward. For instance, in the presence of dish-feed reflections, it will still be feasible to probe only $k_\parallel > 0.15\,h$~Mpc$^{-1}$ with the 58.4~m baselines.

@@

We note that inevitable reflections in the antenna-feed assembly and between reflecting structures across multiple antennas are also significant factors that contaminate the cosmological signal in $k_\parallel$-modes in the {\it EoR window}. Using a novel approach, we provide a formalism to set cosmologically-motivated design requirements directly on the level of suppression required on these reflections in the instrument. This provides a way to set goals on antenna performance and evaluate antenna designs that deliver larger collecting area per antenna.

By accounting for these instrumental systematics, if data is not limited by thermal noise, we demonstrate that HERA should detect EoR in $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ with high significance with a simple spectral weighting even under a conservative analysis criterion that does not involve {\it foreground removal} strategies. All baselines of HERA-19 will not only detect but also help in filtering highly likely EoR models.

We are also investigating advanced strategies for foreground removal that will robustly eliminate these systematics further. Advances in calibration, foreground modeling and subtraction will improve the performance listed here. More optimal analysis techniques and further design improvements will make HERA not only clear the obstacles limiting first-generation experiments but also become transformational for next generation of low-frequency cosmology experiments.

@@

We thank Zaki Ali, Adam Beardsley, and Daniel Jacobs for their valuable inputs during the preparation of this manuscript. This work was supported by the U.S. National Science Foundation (NSF) through awards AST-1440343 and AST-1410719. ARP acknowledges support from NSF CAREER award 1352519. AEW acknowledges support from NSF Graduate Research Fellowship under grant 1122374.
